Predictive modeling with cross validation (this file is a mess)

### 2017-08-26 model fitting/cross_validation.Rmd
model fitting

---
title: "Predictive modeling with cross validation"
author: "Lisa Oshita"
date: "8/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(rms)
```

@@

train1 <- x[splitPlan[[1]]$train, ]
test1 <- x[splitPlan[[1]]$app, ]

train2 <- x[splitPlan[[2]]$train, ]
test2 <- x[splitPlan[[2]]$app, ]

train3 <- x[splitPlan[[3]]$train, ]
test3 <- x[splitPlan[[3]]$app, ]

train4 <- x[splitPlan[[4]]$train, ]
test4 <- x[splitPlan[[4]]$app, ]

train5 <- x[splitPlan[[5]]$train, ]
test5 <- x[splitPlan[[5]]$app, ]
```

@@

fullvars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + pspline(text_length, 3) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3)"
full <- purrr::map2(trains, tests, ~crossval(fullvars, .x, .y))
full
```

@@

```{r}
# turn avg_list into a data frame with rownames???
list <- list(dev, text, tag, newline, devtext, logtext, nospline, lots, lots2, devtag, full)
avg_list <- purrr::map(list, ~get_avgmetrics(.))
avg_list
```

@@

* Martingale residuals
    + plot for each continuous predictor 
    + assess adequacy of functional form of predictors, or suggest a potential form

* Schoenfeld residuals
    + one plot for each predictor 
    + if loess curve is flat, proportional hazard assumption is not violated, doesn't depend on time. Distinct pattern indicates violation of this assumption
    + **from the plots:** new_user, title_questionmark, num_freq_tags (1), ampm (Morning, Night), weekday (Thursday, Saturday), new_user*AMPM (morning, night) may violate PH assumption 
* Score residuals
    + check for influential observations. ID quesitons who's predictor values highly effect parameter estimates
    + plot for each predictor 
* **figure out how to identify certain points on ggplots**

```{r, residuals, echo = FALSE}
full <- coxph(Surv(time_until_answer, answered) ~ strata(new_category) + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + strata(ampm) + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + poly(text_length, 2) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3), data = train1)
# If i fit text_length with poly (orthogonal polynomial), text_length doesn't violate ph anymore 
#=============================================
#martingale residuals
train1$mart <- residuals(full, type = "martingale")
train1$sqrt_score <- sqrt(train1$avg_tag_score)

@@

# looking at points with deviance residuals > |2.5| (suggested in 417 notes)
outliers <- train1[abs(train1$deviance) > 2.5,]
# only 2 of these outliers were not answered

#=============================================
# schoenfeld residuals
schoen <- as.data.frame(residuals(full, type = "schoenfeld"))
# complete event times sorted from shortest to longest
schoen$comp_times <- sort(train1[train1$answered != 0,]$time_until_answer)

plot_schoen <- function(var) {
  ggplot(schoen, aes(x = comp_times, y = schoen[[var]])) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Schoenfeld Residuals for", var)) + 
    scale_y_continuous("Residuals")
}
names <- names(schoen)[-ncol(schoen)]
purrr::map(names, plot_schoen)

@@

#=============================================
# formal test of PH assumption
options(scipen=999)
ph_test <- data.frame(predictors = rownames(cox.zph(full)$table), cox.zph(full)$table) 
rownames(ph_test) = NULL
ph_test <- ph_test %>%
  arrange(p)
ph_test
#=============================================
# score residuals
score <- as.data.frame(residuals(full, type = "score"))

>>>>

---
title: "Predictive modeling with cross validation (this file is a mess)"
author: "Lisa Oshita"
date: "8/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
x <- oshitar::setup()
library(survival); library(stringr); library(rebus); library(dplyr); library(stats); library(rms)
```

@@

train1 <- x[splitPlan[[1]]$train, ]; test1 <- x[splitPlan[[1]]$app, ]

train2 <- x[splitPlan[[2]]$train, ]; test2 <- x[splitPlan[[2]]$app, ]

train3 <- x[splitPlan[[3]]$train, ]; test3 <- x[splitPlan[[3]]$app, ]

train4 <- x[splitPlan[[4]]$train, ]; test4 <- x[splitPlan[[4]]$app, ]

train5 <- x[splitPlan[[5]]$train, ]; test5 <- x[splitPlan[[5]]$app, ]
```

@@

fullvars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + pspline(text_length, 3) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3)"
full <- purrr::map2(trains, tests, ~crossval(fullvars, .x, .y))
full

ugh <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 0, method = 'AIC') + pspline(text_length, df = 0, method = 'AIC') + pspline(avg_tag_length, df = 0, method = 'AIC') + pspline(newline_ratio, df = 0, method = 'AIC'), data = train1)

summary(ugh) 
AIC(ugh, k = 2)
```

@@

```{r}
# model I came to while trying to correct PH assumption violation 
# stratified on ampm, orthogonal polynomial on text_length (new_category still violates PH)
fullvars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + strata(ampm) + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + poly(text_length, 2) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3)"
fullcv <- purrr::map2(trains, tests, ~crossval(fullvars, .x, .y))

# turn avg_list into a data frame with rownames???
list <- list(dev, text, tag, newline, devtext, logtext, nospline, lots, lots2, devtag, full)
avg_list <- purrr::map(list, ~get_avgmetrics(.))
```

@@

```{r, residuals, echo = FALSE}
full <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + strata(ampm) + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + poly(text_length, 2) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3), data = train1)
# If i fit text_length with poly (orthogonal polynomial), text_length doesn't violate ph anymore 
#=============================================
#martingale residuals
# plot for each continuous predictor, assess adequacy of functional form of predictors, or suggest a potential form 
train1$mart <- residuals(full, type = "martingale")
train1$sqrt_score <- sqrt(train1$avg_tag_score)

@@

# looking at points with deviance residuals > |2.5| (suggested in 417 notes)

# only 2 of these outliers were not answered

#=============================================

spline5 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
# schoenfeld residuals
schoen <- as.data.frame(residuals(full, type = "schoenfeld"))
# complete event times sorted from shortest to longest
schoen$comp_times <- sort(train1[train1$answered != 0,]$time_until_answer)

plot_schoen <- function(var) {
  ggplot(schoen, aes(x = comp_times, y = schoen[[var]])) + 
    geom_point() + 
    geom_smooth(method = "loess") + 
    ggtitle(paste("Schoenfeld Residuals for", var)) + 
    scale_y_continuous("Residuals")
}
names <- names(schoen)[-ncol(schoen)]
purrr::map(names, plot_schoen)

@@

#=============================================
# formal test of PH assumption
options(scipen=999)
ph_test <- data.frame(predictors = rownames(cox.zph(spline5)$table), cox.zph(spline5)$table) 
rownames(ph_test) = NULL
ph_test <- ph_test %>%
  arrange(p)

# plots of variables that violated PH assumption in cox.zph test
plot_schoen("new_categoryApple Product")
plot_schoen("new_categoryElectronics")
plot_schoen("new_categoryCamera")
plot_schoen("new_categoryGame Console")
plot_schoen("title_questionmarkTRUE")

#=============================================
# score residuals: check for influential observations. ID quesitons who's predictor values highly effect
score <- as.data.frame(residuals(full, type = "score"))
